Comparing Machine Learning algorithms for classifying cognitive states from fMRI data

### 2012-11-11 article/article.tex
results

\begin{table*}[htpb]
\centering
{\small
\begin{tabular}{|c|c|c|c|c|}
\hline
\textbf{Dimensionality Reduction Method} & \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
PCA & 61.46 &  - & - & - \\
M ROIs & 70.21 & 72.5 & 74.79 & 76.67 \\
Norm-M ROIs & 71.67 & 71.25 & 75.83 & 77.08 \\
Norm-M-SD ROIs & 65.63 & 69.79 & 69.38 & 65.0 \\
Norm/Vox-M ROIs & 73.13 & 76.67 & 76.88 & 79.58 \\
\hline
\end{tabular}}
\caption{Performance results (in \%) of the three classifiers on different methods for feature selection and structure. \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, \emph{SD} for \emph{Standard Deviation}, and \emph{Norm/Vox} for \emph{Normalisation per Voxel index}. Columns marked with an asterisk used the total log-likelihood as a decision function (see section~\ref{sec:trainingMethodsAndClassifiers}).}
\label{tab:results}
\end{table*}

@@

From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used. By using an exact binomial test we found that GNB that used PCA as feature extraction method performed significantly better than chance-level, with a 95\% $CI: 56.9 - 65.8$, $p < 0.001$. However, a Chi-square test for equality of proportions indicated that using PCA as dimentionality reduction method performed significantly worse than a GNB classifier on the means of the selected ROI's, $\chi^2(1) = 7.7848, p < 0.01$. This suggests that maximizing the variance of the data does not maximize information, possibly because the low signal-to-noise ratio of fMRI data does not insure that the principle component with the highest variance also contains a lot of useful information. Additionally, the data may not suited to be aligned by voxel index (see section~\ref{sec:dimensionalityReduction}). The performance listed in Table~\ref{tab:results} is for PCA using 149 components. This value was found by performing classification with a varying amount of components, the results for which are listed in Figure~\ref{fig:pca_components}.

>>>>

\begin{table*}[htpb]
\centering
{\small
\begin{tabular}{|c|c|c|c|c|}
\hline
\textbf{Dimensionality Reduction Method} & \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
PCA & 61.46 &  - & - & - \\
M ROIs & 70.21 & 72.5 & 74.79 & 76.67 \\
Norm-M ROIs & 71.67 & 71.25 & 75.83 & 77.08 \\
Norm-M-SD ROIs & 65.63 & 69.79 & 69.38 & 65.0 \\
Norm/Vox-M ROIs & 73.13 & 76.67 & 76.88 & 79.58 \\
\hline
\end{tabular}}
\caption{Performance results (in \%) of the three classifiers on different methods for feature selection and structure. \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, \emph{SD} for \emph{Standard Deviation}, and \emph{Norm/Vox} for \emph{Normalisation per Voxel index}. Columns marked with an asterisk used the total log-likelihood as a decision function (cf.~Section~\ref{sec:trainingMethodsAndClassifiers}).}
\label{tab:results}
\end{table*}

@@

From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used. By using an exact binomial test we found that GNB that used PCA as feature extraction method performed significantly better than chance-level, with a 95\% $CI: 56.9 - 65.8$, $p < 0.001$. However, a Chi-square test for equality of proportions indicated that using PCA as dimentionality reduction method performed significantly worse than a GNB classifier on the means of the selected ROI's, $\chi^2(1) = 7.7848, p < 0.01$. This suggests that maximizing the variance of the data does not maximize information, possibly because the low signal-to-noise ratio of fMRI data does not insure that the principle component with the highest variance also contains a lot of useful information. Additionally, the data may not suited to be aligned by voxel index (cf.~Section~\ref{sec:dimensionalityReduction}). The performance listed in Table~\ref{tab:results} is for PCA using 149 components. This value was found by performing classification with a varying amount of components, the results for which are listed in Figure~\ref{fig:pca_components}.
